\Title

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
Methods : caption

% The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
% \begin{equation}\label{eq:conv}
% \forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
% \mathcal{C}_\class(\presynaddr,\timev)
% = \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
% \end{equation}
This shows that $\mathcal{C}_\class$ is a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration). Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream. The same type of spatio-temporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}. Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task. By assuming a locally constant velocity, the authors assume the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.

@@

\begin{figure}%[h!]%[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/conv_HD-SNN.pdf}
    % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
    \caption{
    \textbf{- HD-SNN illustration. From left to right:} A 3D representation of the event stream as output of the DVS. A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis. There are as many kernels as the possible number of motion classes in the recording. For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid. For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e. softmax function). The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike. With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class. }
    \label{fig:model}
\end{figure}

@@

We plot here main axis of interests. First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy. This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006}. %\note{we should really do something more serious here...} 
Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018} but also that accuracy is low for a rigid motion which lacks variability.
%
\subsection{Accuracy efficiency trade-off}% for the motion detection task}
%%%-----------------------------------------------------------------

Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection. We observed that the distribution of the kernels' weights is sparse, with most values near zero. As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses. Indeed, the computations are dominated by the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. If the support of information is sparse, then computations can be performed only on those events. Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses. Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.

>>>>

%The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
%\begin{equation}\label{eq:conv}
%\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
%\mathcal{C}_\class(\presynaddr,\timev)
%= \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
%\end{equation}
This shows that $\mathcal{C}_\class$ is a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration). Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream. The same type of spatio-temporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}. Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task. By assuming a locally constant velocity, the authors assume the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.

@@

\begin{figure}%[h!]%[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/conv_HD-SNN.pdf}
    % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
    \caption{
    % \textbf{- HD-SNN illustration. From left to right:} A 3D representation of the event stream as output of the DVS. A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis. There are as many kernels as the possible number of motion classes in the recording. For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid. For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e. softmax function). The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike. With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class. }
    \textbf{Application of HD-SNN to the motion detection task.} \textit{(Left)}~We plot here as  a raster plot a slice of the input event stream (showing ON spikes in red and OFF spikes in blue for each vertical position and time for one given horizontal position in the middle of the sensor), displaying a pure translation motion upwards. A spatio-temporal convolution is applied on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis, and summed over the two polarities. At any given time (represented by the blue dot), we show the extent of the kernels (green dashed frame). \textit{(Middle)}~There are as many kernels as the possible number of motion classes in the recording and we show the activation resulting from the convolution for this specific motion upwards. The activation can be above chance (warm colors) or below chance (cold colors). For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e. softmax function). The output of the MLR gives a probability for each class, associated to a specific kernel. \textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all, we obtain as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class (here red is associated to a motion upwards). Most of the motion is accuratly detected. }
    \label{fig:model}
\end{figure}

@@

We plot here main axis of interests. First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy. This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006}. %\note{we should really do something more serious here...} 
Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018} but also that accuracy is low for a rigid motion which lacks variability.
%
\subsection{Accuracy efficiency trade-off}% for the motion detection task}
%%%-----------------------------------------------------------------
Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection. We observed that the distribution of the kernels' weights is sparse, with most values near zero. As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses. Indeed, the computations are dominated by the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. If the support of information is sparse, then computations can be performed only on those events. Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses. Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.
